Title Goes Here

### 2019-04-29 analysis/paper/paper.Rmd
fixed typos in paper and readme

For example, across two experiments @Gaspelin2016 had participants perform variants of a commonly used  procedure for examining attentional capture---the *additional singleton task* [@Theeuwes1991a; @Theeuwes1992]---while recording eye movements. Eye movements provide a useful index of attention since an eye movement to a location (a saccade) is necessarily preceded by a shift of spatial attention to that location [@Deubel1996]. In this task, participants were required to locate and respond to a shape-singleton target that was presented in an array of non-target shapes. On half of the trials, one of the non-target shapes was rendered in a different colour from the rest of the array so as to be a physically salient colour-singleton distractor. If participants were *more* likely to make a saccade towards the singleton distractor than one of the nonsingleton non-targets, this would indicate that the singleton distractor had captured attention. By contrast, if participants were *less* likely to make a saccade to the singleton distractor than one of the nonsingleton non-targets, this would suggest that attention to the singleton distractor had been suppressed. In @Gaspelin2016\'s Experiment 1, participants were encouraged to adopt a strategy of searching for the "odd-one-out" in the display [i.e, *singleton-search mode*; @Bacon1994], by having the target be randomly either a circle amongst diamonds or a diamond amongst circles on each trial (see Figure \@ref(fig:methodFig)A for an example from the current study). Under these conditions, both stimulus-driven [@Theeuwes1992] and goal-directed [@Bacon1994] accounts of attentional selection would predict capture, as the template that participants use to locate the target (i.e., any singleton) is broad enough to include the singleton distractor. As expected, under these singleton-search conditions Gaspelin et al. found that participants' overt attention was captured by the salient singleton. In Experiment 2, participants were instead encouraged to adopt a more specific target template [i.e., *feature-search mode*; @Bacon1994], by consistently mapping the target identity to one shape (e.g., search for a circle on every trial), and having a heterogeneous set of shapes act as the non-targets (see Figure \@ref(fig:methodFig)B). This manipulation makes searching for any singleton a highly inefficient strategy, as there are multiple singleton nontargets in the display (e.g., a circle, a square, a triangle, and a red shape). Thus, participants should be encouraged to search for the specific features that define the target (i.e., the "circleness" of the circle). Under these conditions, attention to the salient colour-singleton distractor was suppressed relative to the other non-targets, suggesting that the visual attention system treated the salient singleton differently from the other non-targets (inconsistent with goal-directed accounts), yet did not capture gaze (inconsistent with stimulus-driven accounts). By contrast, the signal suppression hypothesis can account for these findings by suggesting that the singleton generated an attentional priority signal as a consequence of its physical salience, but this signal could be suppressed by an inhibitory mechanism under conditions that promoted a more specific search strategy (i.e., feature search). Importantly, this suppression effect was observed even for the most rapidly generated eye movements, which was taken as evidence that the suppression effect operates early in processing, before attention is initially allocated.

@@

## Stimuli
The task was based on that used by @Gaspelin2016 and Pearson et al. [-@Pearson2015; -@Pearson2016]. Each trial consisted of a fixation display, a search display, and a feedback display (see Figure \@ref(fig:methodFig)C). All stimuli were presented on a black background. The fixation display consisted of a white cross (0.5$^\circ$ visual angle), presented in the centre of the screen and surrounded by a white circle (diameter 3.0$^\circ$). The search display consisted of four filled shapes (size 1.6$^\circ$ $\times$ 1.6$^\circ$) presented above, below, left and right of screen centre (eccentricity 4.5$^\circ$). One of the shapes was designated as the target, and the other three shapes were designated non-targets. For all participants, the target was randomly determined to be either a circle or a diamond on each trial [^1]. The identity of the non-target shapes varied by search condition. For the *singleton search* condition, the non-target stimuli were all rendered as the alternative target shape (i.e., if the target was a diamond, all of the non-targets were circles; Figure \@ref(fig:methodFig)B). For the *feature search* condition, the non-target stimuli were randomly selected without replacement from a set of two squares and two hexagons (such that one of the non-targets was always a shape singleton), and were never rendered as the alternative target shape (Figure \@ref(fig:methodFig)C). The target and two of the non-targets were always rendered in grey (CIE $x$, $y$ chromaticity coordinates of .327/.400). On the majority of trials, one of the non-target shapes (the *singleton distractor*) was rendered in either orange or blue (CIE $x$, $y$ coordinates .492/.445 and .192/.216, respectively). On the remaining trials, all shapes were rendered in grey such that there was no singleton distractor. The values of blue and orange had similar luminance (~24.5 cd/m^2^), which was higher than that of grey (~8.3 cd/m^2^). The feedback display showed the points earned on the current trial. If the response time (RT) exceeded the soft-timeout threshold (see below), the message "Too slow" appeared below the reward feedback.

@@

## Design
Participants were randomly assigned to either the singleton search or feature search conditions. For half of the participants in each condition, orange was the high-reward colour and blue was the low-reward colour; these colour-reward relationships were reversed for the remaining participants. There were three types of trial: (1) *high-reward trials*, in which the singleton distractor was rendered in the high-reward colour and 500 points were available for a rapid respnse (see below); (2) *low-reward trials*, in which the singleton distractor was rendered in the low-reward colour and 10 points were available; and (3) *distractor-absent trials*, in which all non-targets were rendered in grey and 10 points were available. The experiment comprised 8 blocks of 60 trials each, for a total of 480 experimental trials. Each block consisted of 20 high-reward trials, 20 low-reward trials, and 20 distractor-absent trials, in random order. Participants took a short break between blocks.

@@

### Effect of reward.
ANOVA analysis of the percentage of reward omissions with factors of trial type (high-reward vs low-reward) and search condition (singleton search vs feature search) revealed a main effect of trial type `r apa_print.afex_aov(omissions_vmac.ANOVA, mse = F, es = "pes")$full_result$dType`. Planned paired-samples *t*-tests confirmed that participants triggered more reward omissions (i.e., were more likely to look at the colour-singleton distractor) on high-reward trials than low-reward trials in both the singleton search condition, `r apa_print(singleton_VMAC_omissions.ttest)$statistic`, $d_z=$ `r singleton_VMAC_omissions.es` (Figure \@ref(fig:OmissionPlot)A) and feature search condition, `r apa_print(feature_VMAC_omissions.ttest)$statistic`, $d_z=$ `r feature_VMAC_omissions.es` (Figure \@ref(fig:OmissionPlot)B), demonstrating an effect of reward on overt attentional capture (i.e., a VMAC effect) regardless of whether participants were encouraged to adopt a more specific target template. Importantly, search condition did not exert a significant main effect, `r apa_print.afex_aov(omissions_vmac.ANOVA, mse = F, es = "pes")$full_result$condition`, or significantly interact with the effect of trial type, $F(1,62) < 0.01$, $p=.985$, $\hat{\eta}^2_p < .0001$. In order to quantify support for the null hypothesis that the influence of reward on attentional capture did not vary depending on participants' target template, we conducted a Bayesian *t*-test to compare the magnitude of the VMAC effect across the two search conditions. This revealed a Bayes Factor of `r apa_print(vmac_condition_bayes)$statistic`, indicating moderate support for the null hypothesis of no difference in the size of the VMAC effect under singleton search and fearure search conditions.

@@

An inconsistency between the results of the current study and those of @Gaspelin2016 comes from the time-course of the attentional capture and suppression effects that were observed. According to the signal suppression hypothesis, suppression is _proactive_---in that it operates prior to the initial allocation of attention---and so suppression should be observed even for the most rapidly generated saccades. Consistent with this idea, @Gaspelin2016 observed an overt attentional suppression effect to physically salient singleton distractors amongst the fastest quartile of saccade latencies. By contrast, in the current experiment we observed overt attentional capture in the fastest quartile of saccade latencies, with the suppression effect emerging on trials with slower saccades. Methodological differences between Gaspelin et al.'s  [-@Gaspelin2016] experiment and the current study may be responsible for this discrepancy. For example, the availability of reward on low-reward trials may have influenced capture, either by somewhat increasing the attentional priority of the stimulus feature associated with low reward, or by otherwise influencing participants' motivation to respond quickly. Alternatively, the discrepancy in findings may reflect our use of two potential targets (circle and diamond) in the feature search condition, as opposed to the same target on all trials in Gaseplin et al.'s procedure. Having two possible targets could reduce the influence of trial-by-trial upweighting of target features [@Maljkovic1994], but may also result in less effective feature search [e.g., @Houtkamp2009; but see @Grubert2015]. Another possible explanation comes from the use of an intermixed trial structure: as participants randomly experienced trials containing either the high- or low-reward colour throughout the course of the task, it is possible that the proactive down-weighting of the salient-but-irrelevant features could not build up enough to prevent capture for the most rapidly initiated saccades [@Gaspelin2018a]. How these methodological differences might influence the time-course of overt attentional capture and suppression remains a question for future research. Nonetheless, the finding does raise questions about whether the attentional suppression effect that we have observed is truly the result of proactive attentional suppression of salient stimuli, as suggested by the signal suppression hypothesis, or potentially the result of rapid covert disengagement from the salient stimuli [see @Theeuwes2010; @VanZoest2004 for similar arguments].

@@

(ref:methodFig) Trial structure of the current task. (A) Participants in the singleton search condition experienced search displays where the target was a unique shape (either a circle amongst diamonds or diamond amongst circles). (B) Participants in the feature search condition experienced search displays where the target was either a circle or diamond shape presented amongst a set of heterogenous shapes. On most trials, one of the non-target shapes was rendered in either orange or blue so as to be a colour singleton distractor, leaving two other nonsingleton non-targets in the display. (C) Participants began by fixating on a centrally presented fixation cross. A search display then appeared, and participants were required to make a rapid saccade to the target. The colour of the singleton distractor indicated the available reward (10 or 500 points), but looking at the distractor led to reward omission.

>>>>

For example, across two experiments @Gaspelin2016 had participants perform variants of a commonly used  procedure for examining attentional capture---the *additional singleton task* [@Theeuwes1991a; @Theeuwes1992]---while recording eye movements. Eye movements provide a useful index of attention since an eye movement to a location (a saccade) is necessarily preceded by a shift of spatial attention to that location [@Deubel1996]. In this task, participants were required to locate and respond to a shape-singleton target that was presented in an array of non-target shapes. On half of the trials, one of the non-target shapes was rendered in a different colour from the rest of the array so as to be a physically salient colour-singleton distractor. If participants were *more* likely to make a saccade towards the singleton distractor than one of the nonsingleton non-targets, this would indicate that the singleton distractor had captured attention. By contrast, if participants were *less* likely to make a saccade to the singleton distractor than one of the nonsingleton non-targets, this would suggest that attention to the singleton distractor had been suppressed. In @Gaspelin2016\'s Experiment 1, participants were encouraged to adopt a strategy of searching for the "odd-one-out" in the display [i.e., *singleton-search mode*; @Bacon1994], by having the target be randomly either a circle amongst diamonds or a diamond amongst circles on each trial (see Figure \@ref(fig:methodFig)A for an example from the current study). Under these conditions, both stimulus-driven [@Theeuwes1992] and goal-directed [@Bacon1994] accounts of attentional selection would predict capture, as the template that participants use to locate the target (i.e., any singleton) is broad enough to include the singleton distractor. As expected, under these singleton-search conditions Gaspelin et al. found that participants' overt attention was captured by the salient singleton. In Experiment 2, participants were instead encouraged to adopt a more specific target template [i.e., *feature-search mode*; @Bacon1994], by consistently mapping the target identity to one shape (e.g., search for a circle on every trial), and having a heterogeneous set of shapes act as the non-targets (see Figure \@ref(fig:methodFig)B). This manipulation makes searching for any singleton a highly inefficient strategy, as there are multiple singleton nontargets in the display (e.g., a circle, a square, a triangle, and a red shape). Thus, participants should be encouraged to search for the specific features that define the target (i.e., the "circleness" of the circle). Under these conditions, attention to the salient colour-singleton distractor was suppressed relative to the other non-targets, suggesting that the visual attention system treated the salient singleton differently from the other non-targets (inconsistent with goal-directed accounts), yet did not capture gaze (inconsistent with stimulus-driven accounts). By contrast, the signal suppression hypothesis can account for these findings by suggesting that the singleton generated an attentional priority signal as a consequence of its physical salience, but this signal could be suppressed by an inhibitory mechanism under conditions that promoted a more specific search strategy (i.e., feature search). Importantly, this suppression effect was observed even for the most rapidly generated eye movements, which was taken as evidence that the suppression effect operates early in processing, before attention is initially allocated.

@@

## Stimuli
The task was based on that used by @Gaspelin2016 and Pearson et al. [-@Pearson2015; -@Pearson2016]. Each trial consisted of a fixation display, a search display, and a feedback display (see Figure \@ref(fig:methodFig)C). All stimuli were presented on a black background. The fixation display consisted of a white cross (0.5$^\circ$ visual angle), presented in the centre of the screen and surrounded by a white circle (diameter 3.0$^\circ$). The search display consisted of four filled shapes (size 1.6$^\circ$ $\times$ 1.6$^\circ$) presented above, below, left and right of screen centre (eccentricity 4.5$^\circ$). One of the shapes was designated as the target, and the other three shapes were designated non-targets. For all participants, the target was randomly determined to be either a circle or a diamond on each trial [^1]. The identity of the non-target shapes varied by search condition. For the *singleton search* condition, the non-target stimuli were all rendered as the alternative target shape (i.e., if the target was a diamond, all of the non-targets were circles; Figure \@ref(fig:methodFig)A). For the *feature search* condition, the non-target stimuli were randomly selected without replacement from a set of two squares and two hexagons (such that one of the non-targets was always a shape singleton), and were never rendered as the alternative target shape (Figure \@ref(fig:methodFig)B). The target and two of the non-targets were always rendered in grey (CIE $x$, $y$ chromaticity coordinates of .327/.400). On the majority of trials, one of the non-target shapes (the *singleton distractor*) was rendered in either orange or blue (CIE $x$, $y$ coordinates .492/.445 and .192/.216, respectively). On the remaining trials, all shapes were rendered in grey such that there was no singleton distractor. The values of blue and orange had similar luminance (~24.5 cd/m^2^), which was higher than that of grey (~8.3 cd/m^2^). The feedback display showed the points earned on the current trial. If the response time (RT) exceeded the soft-timeout threshold (see below), the message "Too slow" appeared below the reward feedback.

@@

## Design
Participants were randomly assigned to either the singleton search or feature search conditions. For half of the participants in each condition, orange was the high-reward colour and blue was the low-reward colour; these colour-reward relationships were reversed for the remaining participants. There were three types of trial: (1) *high-reward trials*, in which the singleton distractor was rendered in the high-reward colour and 500 points were available for a rapid response (see below); (2) *low-reward trials*, in which the singleton distractor was rendered in the low-reward colour and 10 points were available; and (3) *distractor-absent trials*, in which all non-targets were rendered in grey and 10 points were available. The experiment comprised 8 blocks of 60 trials each, for a total of 480 experimental trials. Each block consisted of 20 high-reward trials, 20 low-reward trials, and 20 distractor-absent trials, in random order. Participants took a short break between blocks.

@@

### Effect of reward.
ANOVA analysis of the percentage of reward omissions with factors of trial type (high-reward vs low-reward) and search condition (singleton search vs feature search) revealed a main effect of trial type `r apa_print.afex_aov(omissions_vmac.ANOVA, mse = F, es = "pes")$full_result$dType`. Planned paired-samples *t*-tests confirmed that participants triggered more reward omissions (i.e., were more likely to look at the colour-singleton distractor) on high-reward trials than low-reward trials in both the singleton search condition, `r apa_print(singleton_VMAC_omissions.ttest)$statistic`, $d_z=$ `r singleton_VMAC_omissions.es` (Figure \@ref(fig:OmissionPlot)A) and feature search condition, `r apa_print(feature_VMAC_omissions.ttest)$statistic`, $d_z=$ `r feature_VMAC_omissions.es` (Figure \@ref(fig:OmissionPlot)B), demonstrating an effect of reward on overt attentional capture (i.e., a VMAC effect) regardless of whether participants were encouraged to adopt a more specific target template. Importantly, search condition did not exert a significant main effect, `r apa_print.afex_aov(omissions_vmac.ANOVA, mse = F, es = "pes")$full_result$condition`, or significantly interact with the effect of trial type, $F(1,62) < 0.01$, $p=.985$, $\hat{\eta}^2_p < .0001$. In order to quantify support for the null hypothesis that the influence of reward on attentional capture did not vary depending on participants' target template, we conducted a Bayesian *t*-test to compare the magnitude of the VMAC effect across the two search conditions. This revealed a Bayes Factor of `r apa_print(vmac_condition_bayes)$statistic`, indicating moderate support for the null hypothesis of no difference in the size of the VMAC effect under singleton search and feature search conditions.

@@

An inconsistency between the results of the current study and those of @Gaspelin2016 comes from the time-course of the attentional capture and suppression effects that were observed. According to the signal suppression hypothesis, suppression is _proactive_---in that it operates prior to the initial allocation of attention---and so suppression should be observed even for the most rapidly generated saccades. Consistent with this idea, @Gaspelin2016 observed an overt attentional suppression effect to physically salient singleton distractors amongst the fastest quartile of saccade latencies. By contrast, in the current experiment we observed overt attentional capture in the fastest quartile of saccade latencies, with the suppression effect emerging on trials with slower saccades. Methodological differences between Gaspelin et al.\'s  [-@Gaspelin2016] experiment and the current study may be responsible for this discrepancy. For example, the availability of reward on low-reward trials may have influenced capture, either by somewhat increasing the attentional priority of the stimulus feature associated with low reward, or by otherwise influencing participants\' motivation to respond quickly. Alternatively, the discrepancy in findings may reflect our use of two potential targets (circle and diamond) in the feature search condition, as opposed to the same target on all trials in Gaspelin et al.'s procedure. Having two possible targets could reduce the influence of trial-by-trial upweighting of target features [@Maljkovic1994], but may also result in less effective feature search [e.g., @Houtkamp2009; but see @Grubert2015]. Another possible explanation comes from the use of an intermixed trial structure: as participants randomly experienced trials containing either the high- or low-reward colour throughout the course of the task, it is possible that the proactive down-weighting of the salient-but-irrelevant features could not build up enough to prevent capture for the most rapidly initiated saccades [@Gaspelin2018a]. How these methodological differences might influence the time-course of overt attentional capture and suppression remains a question for future research. Nonetheless, the finding does raise questions about whether the attentional suppression effect that we have observed is truly the result of proactive attentional suppression of salient stimuli, as suggested by the signal suppression hypothesis, or potentially the result of rapid covert disengagement from the salient stimuli [see @Theeuwes2010; @VanZoest2004 for similar arguments].

@@

(ref:methodFig) Trial structure of the current task. (A) Participants in the singleton search condition experienced search displays where the target was a unique shape (either a circle amongst diamonds or diamond amongst circles). (B) Participants in the feature search condition experienced search displays where the target was either a circle or diamond shape presented amongst a set of heterogeneous shapes. On most trials, one of the non-target shapes was rendered in either orange or blue so as to be a colour singleton distractor, leaving two other nonsingleton non-targets in the display. (C) Participants began by fixating on a centrally presented fixation cross. A search display then appeared, and participants were required to make a rapid saccade to the target. The colour of the singleton distractor indicated the available reward (10 or 500 points), but looking at the distractor led to reward omission.
